The role of emotional factors in face processing abilities in Autism Spectrum Conditions

### 2023-07-08 manuscript.tex
Added tables

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The role of emotional factors in face processing abilities in Autism Spectrum Conditions},
  pdfauthor={Natasha Baxter \& Hannah Hobson},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

Table 1 summarises the participants' characteristics.

@@

The Hospital Anxiety and Depression scale (HADS, \protect\hyperlink{ref-zigmond1983a}{Zigmond and Snaith, 1983}) consists of seven questions for anxiety and seven questions for depression. Depression and anxiety scales are scored separately. The HADS has been shown to perform well in measuring anxiety and depression symptoms in non-clinical samples as well as clinical samples (\protect\hyperlink{ref-bjelland2002a}{Bjelland et al., 2002}), including the identification of anxiety and depression in autistic individuals, with good internal consistency and convergent validity for both the anxiety and depression scales (\protect\hyperlink{ref-uljarevic2013a}{Uljarevic and Hamilton, 2013}). The means, standard deviations and distribution of the HADS data for our data from typically developed adults confirm the measure is normally distributed, with suitable variation in scores that will allow for the exploration of the impacts of anxiety and depression on our variables of interest. For HADS anxiety, M = 10.87 (SD = 2.90). For HADS depression, M = 8.69 (SD = 1.94).

@@

Our first sample of non-autistic participants was a sample of 98 participant, (7 of which were considered outliers in our preliminary analyses). We intended to collect at least 100 complete datasets from our autistic sample, and at least 90 datasets that survive removal due to attention check fails, being an outlier, or poor function fit. As noted in our Participants section above, were were unable to achieve this, but osur final sample was 178 participants, after removing incomplete datasets and individuals who failed the attention check. This sample size should still negate any issues with multi-collinearity, and inflation of previous estimate due to publication bias. As previously noted, we included those who self-identify as autistic in our autistic sample in order to aid in recruiting our goal of at least 100 participants. Also, we did not formally checking participants for their clinical diagnosis: as we cannot verify who has a formal diagnosis of autism, we felt it practical to allow people who self-identify as autistic to take part and just ask them if they are or are not formally diagnosed.

@@

@@

In our first series of regressions, a continuous measure of autistic traits, the AQ28, was entered in the final step. Table 3 summarises the results of this regression series. For emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1s and 2 (although the addition of the TAS in step 2 did not lead to a significant increase in the variance explained). The only significant predictor was gender, specifically the dummy variable indicating ``other'' gender. See sub-section ``The role of gender'' for follow up tests on gender effects on emotion PSE. Table 4 summarises the coefficients for the regressions with significant models.

@@

@@

While neither alexithymia nor autistic traits/status were significant predictors of our face processing variables in our analyses, gender did emerge as a significant predictor of emotion PSE scores in the above analyses. Follow-up comparisons comparing emotion PSE between participants who we male, female or selected ``other'' as their gender showed that the other group had significantly lower emotion PSE scores (N = 16, Mdn = .47) compared to both males (N = 74, Mdn = .49) and females (N = 88, Mdn = .50): Kruskal-Wallis H (2) = 9.52, p = .009.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The role of emotional factors in face processing abilities in Autism Spectrum Conditions},
  pdfauthor={Natasha Baxter \& Hannah Hobson},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

Table \ref{tab:table1} summarises the participants' characteristics.

\begin{table}[H]

\caption{\label{tab:table1}Participant Characteristics.}
\centering
\begin{tabular}[t]{llll}
\toprule
 & Full Sample & Autistic Sample & Non-autistic Sample\\
 & (N = 178) & (N = 80) & (N = 98)\\
\midrule
Gender & N Female = 88 & N Female = 48 & N Female = 40\\
 & N Male = 74 & N Male = 21 & N Male = 53\\
 & N Other = 16 & N Other = 11 & N Other = 5\\
\addlinespace
 &  &  \vphantom{5} & \\
Age & 33.14 (13.06) & 41.02 (13.07) & 26.7 (8.88)\\
 &  &  \vphantom{4} & \\
Ethnicity & N Asian = 10 & N Asian = 3 & N Asian = 7\\
 & N Black = 1 & N Black = 1 & N Mixed = 1\\
\addlinespace
 & N Mixed = 4 & N Mixed = 3 & N White = 90\\
 & N White = 163 & N White = 73 & \\
 &  &  \vphantom{3} & \\
HADS-Depression & 8.94 (2.1) & 9.25 (2.12) & 8.68 (2.06)\\
 &  &  \vphantom{2} & \\
\addlinespace
HADS-Anxiety & 9.93 (2.99) & 8.91 (2.76) & 10.77 (2.93)\\
 &  &  \vphantom{1} & \\
AQ-28 & 74.73 (15.31) & 86.96 (11.06) & 64.74 (10.21)\\
 &  &  & \\
TAS-20 & 55.68 (12.86) & 60.88 (12.5) & 51.44 (11.58)\\
\bottomrule
\multicolumn{4}{l}{\textsuperscript{} For non-category data, statistics reported are means with standard}\\
\multicolumn{4}{l}{deviations in parentheses.}\\
\end{tabular}
\end{table}

@@

The Hospital Anxiety and Depression scale (HADS, \protect\hyperlink{ref-zigmond1983a}{Zigmond and Snaith, 1983}) consists of seven questions for anxiety and seven questions for depression. Depression and anxiety scales are scored separately. The HADS has been shown to perform well in measuring anxiety and depression symptoms in non-clinical samples as well as clinical samples (\protect\hyperlink{ref-bjelland2002a}{Bjelland et al., 2002}), including the identification of anxiety and depression in autistic individuals, with good internal consistency and convergent validity for both the anxiety and depression scales (\protect\hyperlink{ref-uljarevic2013a}{Uljarevic and Hamilton, 2013}). The means, standard deviations and distribution of the HADS data for our data from typically developed adults confirm the measure is normally distributed, with suitable variation in scores that will allow for the exploration of the impacts of anxiety and depression on our variables of interest. For HADS anxiety, M = 10.77 (SD = 2.93). For HADS depression, M = 8.68 (SD = 2.06).

@@

Our first sample of non-autistic participants was a sample of 98 participants, (7 of which were considered outliers in our preliminary analyses). We intended to collect at least 100 complete datasets from our autistic sample, and at least 90 datasets that survive removal due to attention check fails, being an outlier, or poor function fit. As noted in our Participants section above, were were unable to achieve this, but our final sample was 178 participants, after removing incomplete datasets and individuals who failed the attention check. This sample size should still negate any issues with multi-collinearity, and inflation of previous estimate due to publication bias. As previously noted, we included those who self-identify as autistic in our autistic sample in order to aid in recruiting our goal of at least 100 participants. Also, we did not formally checking participants for their clinical diagnosis: as we cannot verify who has a formal diagnosis of autism, we felt it practical to allow people who self-identify as autistic to take part and just ask them if they are or are not formally diagnosed.

@@

\begin{table}[H]

\caption{\label{tab:table2}Correlations between alexithymia traits, autistic traits, depression and anxiety symptoms and face processing variables.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{lllllllll}
\toprule
 &  &  &  &  &  &  & Emotion & Identity\\
 & TAS- & AQ & HADS & HADS & Emotion PSE & Identity PSE & attribution & attribution\\
 & TOTAL &  & Depression & Anxiety & (N=174) & (N=176) & (N=174) & (N=170)\\
\midrule
TAS-TOTAL &  & .617 & .076 & -0.308 & 0.023 & 0.115 & 0.093 & 0.042\\
 &  & p<.001 & p=.155 & p<.001 & p=.38 & p=.064 & p=.112 & p=.292\\
\addlinespace
 &  &  &  &  &  &  &  \vphantom{2} & \\
AQ & .617 &  & .149 & -0.326 & 0.033 & 0.049 & 0.137 & 0.106\\
 & p<.001 &  & p=.023 & p<.001 & p=.334 & p=.258 & p=.036 & p=.085\\
 &  &  &  &  &  &  &  \vphantom{1} & \\
HADS Depression & .076 & .149 &  & -0.024 & 0.132 & 0.079 & -0.08 & -0.142\\
\addlinespace
 & p=.155 & p=.023 &  & p=.374 & p=.041 & p=.148 & p=.146 & p=.032\\
 &  &  &  &  &  &  &  & \\
HADS Anxiety & -0.308 & -0.326 & -0.024 &  & 0.024 & 0.004 & -0.032 & -0.019\\
 & p<.001 & p<.001 & p=.374 &  & p=.375 & p=.48 & p=.336 & p=.402\\
\bottomrule
\multicolumn{9}{l}{\textsuperscript{} All correlations are Spearman’s Rank and one-tailed. Note that correlations with the face}\\
\multicolumn{9}{l}{processing variables are with outliers removed, as per Cook’s Distance threshold rules, generated}\\
\multicolumn{9}{l}{for regression series 1 (AQ in final step).}\\
\end{tabular}
\end{table}

@@

In our first series of regressions, a continuous measure of autistic traits, the AQ28, was entered in the final step. Table \ref{tab:table3} summarises the results of this regression series. For emotion attribution thresholds, identity attribution thresholds, and identity PSE, no model was significant at any step. For emotion PSE, the regression model was significant at steps 1s and 2 (although the addition of the TAS in step 2 did not lead to a significant increase in the variance explained). The only significant predictor was gender, specifically the dummy variable indicating ``other'' gender. See sub-section ``The role of gender'' for follow up tests on gender effects on emotion PSE. Table 4 summarises the coefficients for the regressions with significant models.

\begin{table}[H]

\caption{\label{tab:table3}Regression Series 1 (AQ entered in Step 4).}
\centering
\fontsize{5}{7}\selectfont
\begin{tabular}[t]{lllllllll}
\toprule
\multicolumn{1}{l}{} & \multicolumn{2}{l}{Emotion Attribution} & \multicolumn{2}{l}{Identity Attribution} & \multicolumn{2}{l}{Emotion PSE} & \multicolumn{2}{l}{Identity PSE} \\
\multicolumn{1}{l}{} & \multicolumn{2}{l}{Total N (ASC,TD) = 174 (77, 97)} & \multicolumn{2}{l}{Total N (ASC,TD) = 170 (75, 95)} & \multicolumn{2}{l}{Total N (ASC,TD) = 174 (77, 97)} & \multicolumn{2}{l}{Total N (ASC,TD) = 176 (78, 98)} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
 &  & F change, &  & F change, &  & F change, &  & F change,\\
 & F (Df), p-value & p-value & F (Df), p-value & p-value & F (Df), p-value & p-value & F (Df), p-value & p-value\\
\midrule
Model 1 & 1.63 (5,168), .156 & 1.63, .156 & 0.19 (5,164), .968 & 0.19, .968 & 2.4 (5,168), .039 & 2.4, .039 & 1.46 (5,170), .207 & 1.46, .207\\
Model 2 & 1.98 (6,167), .071 & 3.63, .059 & 0.57 (6,163), .753 & 2.49, .116 & 2.34 (6,167), .034 & 1.96, .163 & 1.69 (6,169), .126 & 2.79, .097\\
Model 3 & 1.57 (8,165), .136 & 0.39, .676 & 0.93 (8,161), .49 & 2, .139 & 1.96 (8,165), .055 & 0.83, .439 & 1.26 (8,167), .27 & 0.01, .988\\
\addlinespace
Model 4 & 1.46 (9,164), .165 & 0.62, .432 & 0.93 (9,160), .505 & 0.86, .354 & 1.81 (9,164), .07 & 0.67, .415 & 1.11 (9,166), .356 & 0.03, .868\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[H]

\caption{\label{tab:table4}Coefficients for Emotion PSE Regression Series 1 (AQ entered in Step 4).}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Model &  & Beta & t & p-value\\
\midrule
1 & (Constant) &  & 39.929 & <.001\\
 & Age & -0.119 & -1.588 & .114\\
 & Male gender & -0.134 & -1.717 & .088\\
 & Other gender & -0.218 & -2.792 & .006\\
 & Asian ethnicity & 0.053 & 0.712 & .477\\
\addlinespace
 & Mixed ethnicity & 0.018 & 0.239 & .812\\
2 & (Constant) &  & 23.210 & <.001\\
 & Age & -0.136 & -1.796 & .074\\
 & Male gender & -0.141 & -1.803 & .073\\
 & Other gender & -0.225 & -2.887 & .004\\
\addlinespace
 & Asian ethnicity & 0.061 & 0.809 & .419\\
 & Mixed ethnicity & 0.01 & 0.130 & .896\\
 & TAS-TOTAL & 0.106 & 1.402 & .163\\
3 & (Constant) &  & 13.604 & <.001\\
 & Age & -0.138 & -1.808 & .072\\
\addlinespace
 & Male gender & -0.15 & -1.861 & .065\\
 & Other gender & -0.225 & -2.880 & .005\\
 & Asian ethnicity & 0.049 & 0.652 & .515\\
 & Mixed ethnicity & 0.001 & 0.008 & .993\\
 & TAS-TOTAL & 0.108 & 1.347 & .18\\
\addlinespace
 & HADS Anxiety & 0.026 & 0.322 & .748\\
 & HADS Depression & 0.093 & 1.230 & .22\\
4 & (Constant) &  & 12.636 & <.001\\
 & Age & -0.16 & -1.977 & .05\\
 & Male gender & -0.133 & -1.588 & .114\\
\addlinespace
 & Other gender & -0.232 & -2.945 & .004\\
 & Asian ethnicity & 0.05 & 0.659 & .511\\
 & Mixed ethnicity & 0.004 & 0.055 & .956\\
 & TAS-TOTAL & 0.057 & 0.552 & .581\\
 & HADS Anxiety & 0.035 & 0.424 & .672\\
\addlinespace
 & HADS Depression & 0.083 & 1.084 & .28\\
 & AQ & 0.091 & 0.817 & .415\\
\bottomrule
\end{tabular}
\end{table}

@@

While neither alexithymia nor autistic traits/status were significant predictors of our face processing variables in our analyses, gender did emerge as a significant predictor of emotion PSE scores in the above analyses. Follow-up comparisons comparing emotion PSE between participants who we male, female or selected ``other'' as their gender showed that the other group had significantly lower emotion PSE scores (N = 16, Mdn = 0.47) compared to both males (N = 74, Mdn = 0.49) and females (N = 88, Mdn = 0.5): Kruskal-Wallis H (2) = 9.52, \emph{p} = .009.
